---
title: "NACAP 2022"
author: "Joseph Christian"
date: "February 9, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## library
```{r} 

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))

getwd()
setwd('C:/Users/Christian/OneDrive/Desktop/work from home/NACAP')

figs <- suppressMessages(read_csv('2022FigFrenz_clean.csv'))
```


## variable creation
```{r}
Figs <- figs %>% #complete set of the data
  filter(Totals.State != 'Total')
Figs

popVpart <- Figs %>% #Data for State population vs. ACP participation
  select(Totals.State, '2022.parts', population, `population/participants`)

growth_sqrd <- Figs %>% #Data for State growth rate vs. ACP participant growth rate
  select(Totals.State, `growth rate`, participant_growthRate) 

popVacpGrowth <- Figs %>% #Data for population vs. ACP participant growth rate
  select(Totals.State, population, participant_growthRate) 

grVacpParts <- Figs %>% #Data for State growth rate vs. 2022 ACP participants
  select(Totals.State, `growth rate`, `2022.parts`)

densVprop <- Figs %>% #Data for State population density vs. the proportion of ACP participants to households(gives a look at dependents)
  select(Totals.State, density, `2022.parts`, `2022.households`)

densVacpGrowth <- Figs %>% #Data for State population density vs. ACP growth rate
  select(Totals.State, density, participant_growthRate)

parts2house <- Figs %>% #Data for average dependents per household
  select(Totals.State, `2022.parts`, `2022.households`)

```
## State population vs. ACP participation
```{r}
popVpart #Data for State population vs. ACP participation
popVpart2 <- popVpart %>% 
  arrange(`2022.parts`) %>%
  filter(Totals.State != 'Texas' &
           Totals.State != 'Virginia' &
           Totals.State != 'Florida' &
           Totals.State != 'Washington DC' &
           Totals.State != 'Mississippi') %>%
  mutate(pop2 = population/100000)
popVpart2

popVpart_grf <- ggplot(popVpart2, aes(y = `2022.parts`, x = pop2)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE, fullrange=TRUE) +
  geom_abline(slope = 10,
              intercept = 615.36,
              color = "orange",
              linewidth = 1) +
  scale_x_continuous(limits=c(0, 450), expand = c(0, 0)) +
  scale_y_continuous(limits=c(0, 6000), expand = c(0, 0)) +
  xlab("2022 State Population (by 100,000)") + ylab("2022 Active ACP Participants") +
  theme(panel.border = element_rect(color = 'black',
                                    fill = NA,
                                    linewidth = 2)) +
  ggtitle('2022 Active Participants vs. State Population')

popVpart_grf
cor.test(popVpart2$`2022.parts`, popVpart2$pop2)
```
## State growth rate vs. ACP growth rate
```{r}
growth_sqrd_clean <- growth_sqrd %>% #Data for State growth rate vs. ACP participant growth rate
  rename("growth_rate" = "growth rate") %>%
  drop_na() 

#removing outliers
growth_sqrd_clean2 <- growth_sqrd %>% #Data for State growth rate vs. ACP participant growth rate
  rename("growth_rate" = "growth rate") %>%
  filter(Totals.State != 'New Hampshire' &
           Totals.State != 'Idaho' &
           Totals.State != 'Pennsylvania' &
           Totals.State != 'West Virginia') %>%  #filtering outliers
  drop_na() 
                            
outliers_upper <- quantile(growth_sqrd_clean$growth_rate, 0.999, na.rm = TRUE)  #growth_rate outliers
#outliers_upper <- quantile(growth_sqrd_clean$participant_growthRate, 0.997, na.rm = TRUE)  #participant_growthRate outliers
outliers_upper

outliers_lower <- quantile(growth_sqrd_clean$growth_rate, 0.001, na.rm = TRUE) #growth_rate outliers
#outliers_lower <- quantile(growth_sqrd_clean$participant_growthRate, 0.003, na.rm = TRUE) #participant_growthRate outliers
outliers_lower

#outliers <- which(growth_sqrd_clean$participant_growthRate > outliers_upper | growth_sqrd_clean$participant_growthRate < outliers_lower) #participant_growthRate outliers
outliers <- which(growth_sqrd_clean$growth_rate > outliers_upper | growth_sqrd_clean$growth_rate < outliers_lower)
print(outliers)
#growth_rate outliers
growth_sqrd_clean[24, 1:3] 
growth_sqrd_clean[26, 1:3]
#participant_growthRate
growth_sqrd_clean[3, 1:3] 
growth_sqrd_clean[27, 1:3]

###graph with outliers 
growth_sqrd_grf <- ggplot(growth_sqrd_clean, aes(x=growth_rate, y=participant_growthRate)) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE) +
  ylab("ACP Participant Growth Rate") + xlab("State Population Growth Rate")


###graph with outliers removed
growth_sqrd_grf2 <- ggplot(growth_sqrd_clean2, aes(x=growth_rate, y=participant_growthRate)) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE) +
  ylab("ACP Participant Growth Rate") + xlab("State Population Growth Rate")

##
growth_sqrd_grf
growth_sqrd_grf2
cor.test(growth_sqrd_clean$growth_rate, growth_sqrd_clean$participant_growthRate) #with outliers
```
## State population vs. ACP growth rate
```{r}
popVacpGrowth_clean <- popVacpGrowth %>%
  drop_na()#Data for population vs. ACP participant growth rate

##removing outliers
popVacpGrowth_clean2 <- popVacpGrowth %>%
  filter(Totals.State != 'New Hampshire' &
           Totals.State != 'Pennsylvania') %>%  #filtering outliers)
  drop_na()#Data for population vs. ACP participant growth rate


popVacpGrowth_grf <- ggplot(popVacpGrowth_clean, aes(x=population, y=participant_growthRate)) +
  geom_point() +
  geom_smooth(method=glm, se=FALSE) +
  ylab("ACP Participant Growth Rate") + xlab("State Population")

##without outliers
popVacpGrowth_grf2 <- ggplot(popVacpGrowth_clean2, aes(x=population, y=participant_growthRate)) +
  geom_point() +
  geom_smooth(method=glm, se=FALSE) +
  ylab("ACP Participant Growth Rate") + xlab("State Population")

###graph without outliers
popVacpGrowth_grf2
###graph with outliers
popVacpGrowth_grf

cor.test(popVacpGrowth_clean$population, popVacpGrowth_clean$participant_growthRate)
cor.test(popVacpGrowth_clean2$population, popVacpGrowth_clean2$participant_growthRate)
```
## State growth rate vs. ACP participation
```{r}
grVacpParts_clean <- grVacpParts %>%   #Data for State growth rate vs. 2022 ACP participants
  rename("growth_rate" = "growth rate") %>%
  drop_na()

##removing outliers
grVacpParts_clean2 <- grVacpParts %>%   #Data for State growth rate vs. 2022 ACP participants
  rename("growth_rate" = "growth rate") %>%
  filter(Totals.State != 'Idaho' &
           Totals.State != 'West Virginia') %>%  #filtering outliers
  drop_na()

grVacpParts_grf <- ggplot(grVacpParts_clean, aes(x=growth_rate, y=`2022.parts`)) +
  geom_point() +
  geom_smooth(method=glm, se=FALSE) +
  ylab("2022 ACP Participants") + xlab("State Population Growth Rate")


grVacpParts_grf2 <- ggplot(grVacpParts_clean2, aes(x=growth_rate, y=`2022.parts`)) +
  geom_point() +
  geom_smooth(method=glm, se=FALSE) +
  ylab("2022 ACP Participants") + xlab("State Population Growth Rate")

###graph with outliers included
grVacpParts_grf
###graph without outliers
grVacpParts_grf2

cor.test(grVacpParts_clean$growth_rate, grVacpParts_clean$`2022.parts`)
cor.test(grVacpParts_clean2$growth_rate, grVacpParts_clean2$`2022.parts`)
```
## State population density vs. (participants/households)
```{r}
densVprop_clean <- densVprop %>% #Data for State population density vs. the proportion of ACP participants to households(gives a look at dependents)
  drop_na() %>%
  mutate('fullHouse' = `2022.households`/`2022.parts`) %>%
  arrange(density)

densVprop_grf <- ggplot(densVprop_clean, aes(x=density, y=fullHouse)) +
  geom_point() +
  geom_smooth(method=glm, se=FALSE) +
  ylab("2022 ACP Residents/House") + xlab("State Density")

densVprop_grf
cor.test(densVprop_clean$density, densVprop_clean$fullHouse)
```
## State population density vs. ACP growth rate
```{r}
#densVacpGrowth #Data for State population density vs. ACP growth rate

densVacpGrowth_clean <- densVacpGrowth %>% #Data for State population density vs. the proportion of ACP participants to households(gives a look at dependents)
  drop_na() %>%
  arrange(density)

##removing outliers
densVacpGrowth_clean2 <- densVacpGrowth %>% #Data for State population density vs. the proportion of ACP participants to households(gives a look at dependents)
  filter(Totals.State != 'New Hampshire' &
           Totals.State != 'Pennsylvania') %>% 
  drop_na() %>%
  arrange(density)

densVprop_grf <- ggplot(densVacpGrowth_clean, aes(x=density, y=participant_growthRate)) +
  geom_point() +
  geom_smooth(method=glm, se=FALSE) +
  ylab("2022 ACP Growth Rate") + xlab("State Density")

##outliers removed
densVprop_grf2 <- ggplot(densVacpGrowth_clean2, aes(x=density, y=participant_growthRate)) +
  geom_point() +
  geom_smooth(method=glm, se=FALSE) +
  ylab("2022 ACP Growth Rate") + xlab("State Density")

densVprop_grf
densVprop_grf2
densVprop_clean
cor.test(densVacpGrowth_clean$density, densVacpGrowth_clean$participant_growthRate)
cor.test(densVacpGrowth_clean2$density, densVacpGrowth_clean2$participant_growthRate)
```
## ACP participants vs. ACP households
```{r}
#parts2house #Data for average dependents per household
parts2house_clean <- parts2house %>% 
  drop_na()

parts2house_grf <- ggplot(parts2house_clean, aes(x=`2022.parts`, y=`2022.households`)) +
  geom_point() +
  geom_smooth(method=glm, se=FALSE) +
  xlab("2022 ACP Participation") + ylab("2022 ACP Households")

parts2house_grf2 <- ggplot(parts2house_clean, aes(x=`2022.parts`, y=`2022.households`)) +
  geom_point() +
  geom_smooth(method=loess, se=FALSE) +
  xlab("2022 ACP Participation") + ylab("2022 ACP Households")


parts2house_grf
parts2house_grf2
cor.test(parts2house_clean$`2022.parts`,parts2house_clean$`2022.households`)
```
```{r}
brenda <- Figs %>% 
  transmute('perCapita' = `population/participants`, 'State' = Totals.State) %>% 
  drop_na() %>%
  arrange(perCapita)
brenda
brend_grfz <- ggplot(brenda, aes(x=reorder(State, -perCapita), y=perCapita)) +
  geom_col() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
brend_grfz
```
```{r}
agg <- Figs %>% 
  select(Totals.State, `2022.parts`, `2021.parts`, `2022.households`, `2021.households`, `2022.totals`, `2021.totals`) %>% 
  drop_na()
participants2022 <- mean(agg$`2022.parts`)
participants2021 <- mean(agg$`2021.parts`)
households2022 <- mean(agg$`2022.households`)
households2021 <- mean(agg$`2021.households`)
totals2022 <- mean(agg$`2022.totals`)
totals2021 <- mean(agg$`2021.totals`)
participants2022
participants2021
households2022
households2021
totals2022
totals2021

```
```{r}
avgs <- suppressMessages(read_csv('2022FigFrenz_clean_agg.csv'))
```
```{r}
avgs_grf <- ggplot(avgs, aes(x = group, y = average, fill = (as.factor(year)))) +
  geom_bar(width = 0.5, 
           position = position_dodge(0.6), 
           stat = 'identity', 
           color = 'black') +
  geom_text(aes(label = average), 
            position = position_dodge(0.6), 
            vjust = -0.3) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank(),
        panel.border = element_rect(color = 'black',
                                    fill = NA,
                                    linewidth = 1.5)) +
  ggtitle('Program Averages') +
  scale_x_discrete(labels = c('Participants','Households','Total Served'))
avgs_grf
```





